Exploring Bounds

### 2015-04-23 draft.tex
progress maybe

@@

Note that $n_{i,\alpha}(M - n_{i,\alpha})$ is multiplying the number of times $\alpha$ occurs in column $i$ by the number of times it doesn't occur in column $i$.  In other words, we're counting the number of times two elements of the same
column are not equivalent, exactly the condition that constitutes our Hamming distance.   By summing over all $\alpha$ and then all columns, we arrive at the same value generated by \ref{plotkin_1}.

@@

It is also interesting to consider a more broader picture of what is going on here.  Below is a graph comparing rates of codes versus the Sphere Packing and Plotkin bounds
that we can calculate for those values for different values of q.

@@

\section{Lexicodes}
We end our discussion on bounds with an interesting class of binary linear codes
called lexicodes, short for lexicographic codes.  Lexicodes are linear and meet the Gilbert bound; we will discuss the former and prove the latter.  Note that we are 
working in $\mathbb{F}_2$.

We construct the lexicode $\mathcal{L}$ of length $n$ and minimum distance $d$ using a basic greedy algorithm as follows.  First, order all $n$-tuples in lexicographic order:
$0\cdots000$, $0\cdots001$, $0\cdots010$, $0\cdots011$, $0\cdots100$, et cetera.  Put the zero vector $\mathbf{0}$ in $\mathcal{L}$.  Next, find the first vector $\mathbf{x}$
of weight $d$ in the lexicographic ordering, and put it in $\mathcal{L}$.  Then, find the next vector in  the lexicographic ordering whose distance from each vector in $\mathcal{L}$ is 
$d$ or more and add it to $\mathcal{L}$, repeating this process until you have used up your lexicographic list.

@@

\begin{exmp}
Consider the codes $\mathcal{L}_i$ of length $5$ and minimum distance $2$.  First, we must construct the entire lexicode $\mathcal{L}$.  To do so, we enumerate
all 32 vectors of $\mathbb{F}_2^5$ in lexicographic order: $00000$,$00001$,$00010$,$\ldots$,$11111$.  We find that $\mathcal{L} = \{00000, 00011,00101,00110,01001,
01010,01100,01111,10001,10010,\\ 10100,10111,11000,11011,11101,11110\}$.  As a sanity check, note that $\left | \mathcal{L} \right | = 16$, a power of 2.  Consequently, we have
$\mathcal{L}_1 = \{00000,00011\}$, $\mathcal{L}_2 = \{00000, 00011,00101,\\ 00110\}$, $\mathcal{L}_3 = \{00000, 00011,00101,00110,01001,01010,01100,01111\}$, and
$\mathcal{L}_4 = \mathcal{L}$. We can check to make sure that each $\mathbf{c}_j$ is a linear combination of the $\mathbf{c}_{2^i}$s. Consider $\mathcal{L}_3$:

@@

Note that the greedy algorithm we descibed in the beginning of the section requires a strictly lexigraphical ordering, or else the code that is produced will not necessarily
be linear.  Again consider the case of $\mathbb{F}_2^5$, but instead of the lexigraphical ordering, shift every vector up one such that the ordering is identical except now the
zero vector is at the bottom of the list.  If we run our algorithm on this ordering, we will clearly not choose the zero vector to be a part of $\mathcal{L}$ and hence 
$\mathcal{L}$ will not be linear.

@@

\begin{proof}
Consider an arbitrary lexicode $\mathcal{L}$.  Recall our construction of $\mathcal{L}$ with vectors in $\mathbb{F}_q^n$, wherein we visit each vector in our ordering,
and if it is distance $d$ or more to every current element in $\mathcal{L}$ then we add it to our set.  More importantly, if a vector is distance $d-1$ or less to any vector in
$\mathcal{L}$, then it is not selected.  Since we scan over every possible vector in $\mathbb{F}_q^n$, this implies that all vectors in $\mathbb{F}_q^n$ are distance less than
or equal to $d-1$ from any vector in $\mathcal{L}$, as if at any point of our algorithm we reached some $v \in \mathbb{F}_q^n$, $v \not\in \mathcal{L}$,  we would have then
added it $\mathcal{L}$, a contradiction.  Hence the covering radius of $\mathcal{L}$ is $d-1$.

Now, it is enough to show that any code $\mathcal{C}$ with covering radius $d-1$ or less meets the Gilbert bound.  Recall that the Gilbert bound can be thought of as a restatement of the 
following fact: (number of spheres around codewords) $\cdot$ (number of vectors in each sphere) $\ge$ (number of vectors in $\mathbb{F}_q^n$).  Since our covering radius
is $d-1$ or less, none of the spheres for our lexicodes will overlap, and thus there will be no double counting of vectors when counting up the vectors in each individual sphere.
Hence, $\mathcal{C}$ meets the Gilbert bound.
\end{proof}

Consequently, for our asymptotic bounds, we are interested in the largest possible rate for a family of codes over $\mathbb{F}_q$ of lengths going to infinity with relative 
distances approaching some constant $\delta$.  In other words, we consider the equation:

[What does this imply?]

[explain what this means, what are the implications]

\begin{theorem}
If $0 \le \delta \le 1$, then $\alpha_q(\delta) \le 1 - \delta$.
\end{theorem} 
This theorem follows directly from the Singleton bound.

>>>>

\newtheorem{lemma}{Lemma}
\numberwithin{lemma}{subsection}

@@

Note that $n_{i,\alpha}(M - n_{i,\alpha})$ is multiplying the number of times $\alpha$ occurs in column $i$ by the number of times it doesn't occur in column $i$.  In other words, we're counting the number of times two elements of the same
column are not equivalent, exactly the condition that constitutes the Hamming distance.   By summing over all $\alpha$ and then all columns, we arrive at the same value generated by \ref{plotkin_1}.

@@

Using MATLAB, I sought to visualize how the Plotkin and Sphere Packing bounds compare.  To do so, I plotted their values over $8\le n \le 25$, with $d>rn$.

In order to turn my discrete 3d data into a smooth mesh surface, I needed a mechanism to smooth out my data.  My solution was to utilize Delaunay triangulation.

It is also interesting to consider a broader picture of how Plotkin compares to Sphere Packing, Greismer, and other bounds we have considered in lecture.  Below is a graph comparing rates of codes versus the Sphere Packing and Plotkin bounds
that we can calculate for those values for different values of q.

@@

\subsection{Elias Upper Bound}

Another upper bound on $A_q(n,d)$ is due to Elias in 1960.  The Elias bound is generally weaker than other bounds we have discussed, but is important because of the asymptotic
form of this bound that we will discuss towards the end of this paper.

In order to prove the Elias bound, two lemmas are necessary.

\begin{lemma}
Let $\mathcal{C}$ be an $(n,K,d)$ code over $\mathbb{F}_q$ such that all codewaords have weights at most w, where $w \le rn$ with $r=1-q^{-1}$.  Then
\begin{equation}
d\le \frac{Kw}{K-1}\left(2- \frac{w}{rn}\right).
\end{equation}
\end{lemma}

\begin{lemma}
Suppose C is an (n,M,d) code over $\mathbb{F}_q$.  Then there is an $(n,M,d)$ code $\mathcal{C}'$ over $\mathbb{F}_q$ with an (n,K,d) subcode $\mathcal{A}$ containing
only codewords of weight at most w such that $K \ge M V_q(n,w) / q^n$.
\end{lemma}

\begin{theorem}
Let $r=  1-q^{-1}$.  Suppose that $w\le rn$ and $w^2 - 2rnw + rnd > 0$.  Then

\begin{equation}
A_q(n,d) \le \frac{rnd}{w^2 - 2rnw + rnd}\cdot \frac{q^n}{V_q(n,w)}.
\end{equation}
\end{theorem}

We can work through a few examples to see how the Elias bound operates.

\begin{exmp}
$n=14$, $d=6$, $q=2$. 
Recall from before that $A_2(n,d)=A_2(n-1,d-1)$ if $d$ is even.  We will again use this fact to find the best possible bound that this theorem can provide.

 For $w=0$, 
we get $A_2(13,5)\le \frac{.5\cdot13\cdot5}{.5\cdot13\cdot5}\cdot\frac{2^{13}}{V_q(13,0)} = 2^{13} = 8192$; similarly, $A_2(14,6) = 2^{14} = 16384$.

 For $w=1$, 
we get $A_2(13,5)\le \frac{.5\cdot13\cdot5}{1 - 2\cdot.5\cdot13 + .5\cdot13\cdot5}\cdot\frac{2^{13}}{V_q(13,1)} = \frac{32.5}{1-13+32.5}\cdot\frac{2^{13}}{1+13} = 927$;
similarly $A_2(14,6)\le \frac{.5\cdot14\cdot6}{1 - 2\cdot.5\cdot14 + .5\cdot14\cdot6}\cdot\frac{2^{14}}{V_q(14,1)} = \frac{42}{1-14+42}\cdot\frac{2^{14}}{1+14} = 1581$;

Consequently, for our asymptotic bounds, we are interested in the largest
possible rate for a family of codes over $\mathbb{F}_q$ of lengths going to
infinity with a relative distance of some constant $\delta$.  In other words, we consider the equation:

To better understand what is going on here, consider the case of an $[n,k,d]$ linear code $\mathcal{C}$.  Before, we wanted to know how large we could get $k$ given
$n$ and $d$.  Now, we want to know how large we can get $\frac{k}{n}$ given $\frac{d}{n}=\delta$ as $n\to\infty$.  This idea generalizes to non-linear codes as 
in the non-asymptotic cases we discussed in class.

For the rest of this section, we will consider some bounds on $A_q(n,\delta n)$

The asymptotic bound follows almost directly from the non-asymptotic variation.

\begin{theorem}
If $0 \le \delta \le 1$, then $\alpha_q(\delta) \le 1 - \delta$.
\end{theorem}

\begin{exmp}

\section{Lexicodes}
We end our discussion on bounds with an interesting class of binary linear codes
called lexicodes, short for lexicographic codes.  Lexicodes are linear and meet the Gilbert bound; we will discuss the former and prove the latter.  Note that for this discussion we are 
solely working in $\mathbb{F}_2$.

We construct the lexicode $\mathcal{L}$ of length $n$ and minimum distance $d$ using a basic greedy algorithm as follows.  First, order all $n$-tuples in an array $A$ in lexicographic order:
$A[0] = 0\cdots000$, $A[1]=0\cdots001$, $A[2]0\cdots010$, $A[3]=0\cdots011$, \ldots, $A[2^n-1] = 1\cdots111$.  Put the first vector from  $A$ (always the $\mathbf{0}$ vector) in $\mathcal{L}$.  Next, find the first vector $\mathbf{x}$
of weight $d$ in the lexicographic ordering, and put it in $\mathcal{L}$.  Then, find the next vector in the lexicographic ordering in $A$ whose distance from each vector in $\mathcal{L}$ is 
$d$ or more and add it to $\mathcal{L}$.  Repeat this process until you have scanned through all of $A$ once.

@@

\begin{exmp}
Consider the codes $\mathcal{L}_i$ of length $5$ and minimum distance $2$.  First, we must construct the entire lexicode $\mathcal{L}$.  To do so, we enumerate
all 32 vectors of $\mathbb{F}_2^5$ in lexicographic order: $00000$,$00001$,$00010$,$\ldots$,$11111$.  We find that $\mathcal{L} = \{00000, 00011,00101,00110,01001,
01010,01100,01111,10001,10010,\\ 10100,10111,11000,11011,11101,11110\}$.  As a sanity check, note that $\left | \mathcal{L} \right | = 16$, a power of $q$.  Consequently, we have
$\mathcal{L}_1 = \{00000,00011\}$, $\mathcal{L}_2 = \{00000, 00011,00101,\\ 00110\}$, $\mathcal{L}_3 = \{00000, 00011,00101,00110,01001,01010,01100,01111\}$, and
$\mathcal{L}_4 = \mathcal{L}$. We can check to make sure that each $\mathbf{c}_j$ is a linear combination of the $\mathbf{c}_{2^i}$s. For this example, consider $\mathcal{L}_3$:

Notice that this is simply the even weight vectors.  In fact, it is true that for any $q$

THIS IS THE EVEN WEIGHT CODE! prove it.

@@

Note that the greedy algorithm we descibed in the beginning of the section requires a strictly lexigraphical ordering, or else the code that is produced will not necessarily
be linear.  Again consider the case of $\mathbb{F}_2^5$, but instead of the lexigraphical ordering, shift every vector up one such that the ordering is identical to before, except now the
zero vector is at the bottom of the list.  If we run our algorithm on this ordering, we will clearly not choose the zero vector to be a part of $\mathcal{L}$ and hence 
$\mathcal{L}$ will not be linear.

@@

\begin{proof}
Consider an arbitrary lexicode $\mathcal{L}$.  Recall our construction of $\mathcal{L}$ with vectors in $\mathbb{F}_q^n$, wherein we visit each vector in our ordering,
and if it is distance $d$ or more to every current element in $\mathcal{L}$ then we add it to our set.  At the same time, notice that when a vector is visited in our algorithm, 
if it has distance $d-1$ or less to any vector in $\mathcal{L}$, then it is not placed in $\mathcal{L}$.  Since all vectors in $\mathbb{F}_q^n$ are visited in our algorithm, any
$v \in \mathbb{F}_q^n$ must either be in $\mathcal{L}$ or must have distance $d-1$ or less to some vector in $\mathcal{L}$.  Hence the covering radius of $\mathcal{L}$ is $d-1$.

Consequently, it is enough to show that any code $\mathcal{C}$ with covering radius $d-1$ or less meets the Gilbert bound.  Recall that the Gilbert bound can be thought of as a restatement of the 
following fact: (number of spheres) $\cdot$ (number of vectors in a sphere with radius $d-1$) $\ge$ (number of vectors in $\mathbb{F}_q^n$).  Since our covering radius
is $d-1$ or less, and we have that all codewords are distance $d$ or more apart, consequently no sphere of radius $d-1$ in $\mathbb{F}_q^n$ can contain more than one codeword.
As a result, 
Hence, $\mathcal{C}$ meets the Gilbert bound.
\end{proof}

Some notable codes we have discussed in class are in fact lexicodes. For example, the binary lexicodes with $d$=3 and $n=2^m-1$ are the binary hamming codes.  It can be quite
tedious to compute and confirm something like this by hand, so I wrote a short C++ program that implements the Lexicode algorithm as described above.

\begin{exmp}
Let $n=7$, $d=3$.  We find $\mathcal{L}= \{0000000,000111,  \}$
\end{exmp}

Need to talk about paper
http://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1057187
